### 2013-11-29 paper/main.tex
documented delta to steady state conversion
The transfer function to convert delta values to
steady state has been derived and documented.

\usepackage{tikz}
\usetikzlibrary{calc,arrows,positioning,shadows}
% control system tikz objects
\tikzset{shadow/.style={drop shadow,fill=white}}
\tikzset{sum/.style={circle,draw,very thick,shadow,minimum size=6mm}}
\tikzset{sumnp/.style={sum,
				label={{above left,xshift=1.0mm,yshift=-1.0mm}:$+$},
				label={{below left,xshift=1.0mm,yshift=1.0mm}:$-$}} }
\tikzset{sumpn/.style={sum,
				label={{above left,xshift=1.0mm,yshift=-1.0mm}:$+$},
				label={{above right,xshift=-1.0mm,yshift=-1.0mm}:$-$}} }
\tikzset{sumxpxp/.style={sum,
				label={{above left,xshift=1.0mm,yshift=-1.0mm}:$+$},
				label={{above right,xshift=-1.0mm,yshift=-1.0mm}:$+$}} }
\tikzset{gain/.style={rectangle,draw,very thick,inner sep=2mm,shadow}}

@@

\end{center}
\caption{Linear engine model.}\label{fig:lem}
\end{figure}

@@

This model has one complication that makes it difficult to use compared
to typical control systems.
All the inputs and outputs are specified as deltas ($\Delta$).
The torque input, for example, cannot be given as a constant ($T$) because this
actually means $\Delta T$

To overcome this complication a simple feedback controller can be used
as show in Figure \ref{fig:scont}.
And the open loop controller, with a fixed input duty cycle, is shown
in Figure \ref{fig:openloop}.
Notice that in this example the torque has transient characteristics and
varies over time.
The output of this system is show in Figure \ref{fig:olplot}.

\clearpage
\begin{figure}[!htbp]
\begin{center}

\begin{tikzpicture}[>=triangle 60]
\node[gain] (gain1) [] {$1 - z^{-1}$};
\draw [->] (-30mm,0) node[above,yshift=1mm]{$X$} -- (gain1.west);
\draw [->] (gain1.east) -- (30mm,0) node[above,yshift=1mm]{$\Delta X$};
\end{tikzpicture}

\end{center}
\caption{$Z$ transform used to accumulate the input and convert
a steady state input to delta output.
For its derivation, see Appendix \ref{app:cdelta}.}
\label{fig:scont}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[scale=0.5]{img/schematic-no_control-ed1}
\end{center}
\caption{Linear engine model with an open loop controller.
Inputs go through the simple controller to provide steps to the
engine model.
The output rpm, since it is in steps ($\Delta N$), is accumulated.}
\label{fig:openloop}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.8]{img/linear_engine_model_no_control_plot}
\end{center}
\caption{Output of linear engine controller with a transient torque
input and a fixed input duty cycle.  Matlab source given in
Appendix \ref{app:olplot}}
\label{fig:olplot}
\end{figure}

@@

% {{{ Steady State to Delta $Z$ Transform Derivation
\clearpage
\section{Steady State to Delta $Z$ Transform Derivation}
\label{app:cdelta}

@@

\begin{tikzpicture}[>=triangle 60]
\node[gain] (gain1) [] {$1 - z^{-1}$};
\draw [->] (-30mm,0) node[above,yshift=1mm]{$u$} -- (gain1.west);
\draw [->] (gain1.east) -- (30mm,0) node[above,yshift=1mm]{$y$};
\end{tikzpicture}

@@

>>>>

\usepackage{tikz}
\usetikzlibrary{calc,arrows,positioning,shadows}
%
% control system tikz objects
\tikzset{shadow/.style={drop shadow,fill=white}}
\tikzset{sum/.style={circle,draw,very thick,shadow,minimum size=6mm}}
\tikzset{sumnp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{below right,xshift=-1.0mm,yshift=0.0mm}:$-$}} }
\tikzset{sumpp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{below right,xshift=-1.0mm,yshift=0.0mm}:$+$}} }
\tikzset{sumpn/.style={sum,
				label={{below left,xshift=0.0mm,yshift=1.0mm}:$+$},
				label={{above right,xshift=-1.0mm,yshift=0.0mm}:$-$}} }
\tikzset{sumxpxp/.style={sum,
				label={{above left,xshift=0.0mm,yshift=-1.0mm}:$+$},
				label={{above right,xshift=-0.0mm,yshift=-1.0mm}:$+$}} }
\tikzset{gain/.style={rectangle,draw,very thick,inner sep=2mm,shadow}}

@@

\end{center}
\caption{Linear engine model of a modern 4.6L V-8.}\label{fig:lem}
\end{figure}

@@

Because all the inputs and outputs are defined as deltas ($\Delta$)
this model cannot be used directly with typical control systems
which use steady state values.
It is possible convert these deltas to steady state equivalents.
Figure \ref{fig:sdtf} shows the transfer function to convert steady
state values to delta values.
Figure \ref{fig:dstf} shows the transfer function to delta values
to steady state values.

% {{{ fig:sdtf
\begin{figure}[!htbp]

\begin{center}
\begin{tikzpicture}[>=triangle 60]
\node[gain] (gain1) [] {$1 - z^{-1}$};
\draw [->] (-20mm,0) node[above,yshift=1mm]{$X$} -- (gain1.west);
\draw [->] (gain1.east) -- (20mm,0) node[above,yshift=1mm]{$\Delta X$};
\end{tikzpicture}
\end{center}

\caption{The $Z$ transform used to accumulate the input and convert
a steady state input to delta output.
Its derivation is given in Appendix \ref{app:cdelta}.}
\label{fig:sdtf}

\end{figure}
% }}}

% {{{ fig:dstf
\begin{figure}[!htbp]

\begin{center}
\begin{tikzpicture}[>=triangle 60]
\node[gain] (gain1) [] {$\dfrac{1}{1 - z^{-1}}$};
\draw [->] (-20mm,0) node[above,yshift=1mm]{$\Delta X$} -- (gain1.west);
\draw [->] (gain1.east) -- (20mm,0) node[above,yshift=1mm]{$X$};
\end{tikzpicture}
\end{center}

\caption{The $Z$ transform used to convert delta input to a steady
state output.
Its derivation is given in Appendix \ref{app:dsderiv}.}
\label{fig:dstf}

\end{figure}
% }}}

% TODO - delta to steady state

Typical control systems have an associated time step which can be
varied and is crucial in determining the performance of
the controller with regard to the Nyquist frequency.
This model has no such constraint because one calculation step
corresponds directly to one engine ignition event.
The controller is effectively running at the same ``clock speed'' as
the engine.

% TODO - steady state model

%And the open loop controller, with a fixed input duty cycle, is shown
%in Figure \ref{fig:openloop}.
%Notice that in this example the torque has transient characteristics and
%varies over time.
%The output of this system is show in Figure \ref{fig:olplot}.

% TODO - open loop performance

%\begin{figure}[!htbp]
%\begin{center}
%\includegraphics[scale=0.5]{img/schematic-no_control-ed1}
%\end{center}
%\caption{Linear engine model with an open loop controller.
%Inputs go through the simple controller to provide steps to the
%engine model.
%The output rpm, since it is in steps ($\Delta N$), is accumulated.}
%\label{fig:openloop}
%\end{figure}

%\begin{figure}
%\begin{center}
%\includegraphics[scale=0.8]{img/linear_engine_model_no_control_plot}
%\end{center}
%%\caption{Output of linear engine controller with a transient torque
%input and a fixed input duty cycle.  Matlab source given in
%Appendix \ref{app:olplot}}
%\label{fig:olplot}
%\end{figure}

@@

% {{{ Steady State to Delta Transform Derivation
\clearpage
\section{Steady State to Delta Transform Derivation}
\label{app:cdelta}

@@

\begin{tikzpicture}[>=triangle 60]
\node[gain] (gain1) [] {$1 - z^{-1}$};
\draw [->] (-20mm,0) node[above,yshift=1mm]{$u$} -- (gain1.west);
\draw [->] (gain1.east) -- (20mm,0) node[above,yshift=1mm]{$y$};
\end{tikzpicture}

@@

% {{{ Delta to Steady State Transform Derivation
\clearpage
\section{Delta to Steady State Transform Derivation}
\label{app:dsderiv}

To convert a delta input to a steady state output it should
sum the history of values.
Figure \ref{fig:dc1} shows the system.

% {{{ fig:dc1
\begin{figure}[hpb!]
\begin{center}

	\node[sumpp]  	(sum1)	[] 						{};
	\node[gain] 	(gain1)	[below right=of sum1]	{$z^{-1}$};

	\draw [->] (gain1.west) -| (sum1.south) node[left,yshift=-10mm]{$r$};
	\draw [<-] (sum1.west) -- ++(-20mm,0) node[above]{$u$};
	\draw [->] (sum1.east) -- (50mm,0) node[above]{$y$};
	\draw [->] ($ (sum1.east) + (30mm,0) $)
			|- (gain1.east);
\end{tikzpicture}

\caption{System to convert delta input to steady state output.}
\label{fig:dc1}
\end{figure}
% }}}

This system can be simplified in to a single transfer function
as given by Equation \ref{eq:dc2s} and shown in Figure \ref{fig:dc1s}.

\begin{align}
	y &= u + r \notag \\
	r &= y \cdot z^{-1} \notag \\
	y &= u + y z^{-1} \notag \\
	y &= y(1 - z^{-1}) \notag \\
	\Aboxed{ \frac{y}{u} &= \frac{1}{1 - z^{-1}} } \label{eq:dc2s}
\end{align}

% {{{ fig:dc1s
\begin{figure}[!htbp]
\begin{center}

\begin{tikzpicture}[>=triangle 60]
\node[gain] (gain1) [] {$1 - z^{-1}$};
\draw [->] (-20mm,0) node[above,yshift=1mm]{$u$} -- (gain1.west);
\draw [->] (gain1.east) -- (20mm,0) node[above,yshift=1mm]{$y$};
\end{tikzpicture}

\end{center}
\caption{Simplified system to convert a delta input to
a steady state output.}
\label{fig:dc1s}
\end{figure}
% }}}

Figure \ref{fig:dc2_plot} shows the response of this system given
an arbitrary input
\footnote{This input is actually the output of the steady state input
to delta output given in Appendix \ref{app:cdelta}.
They are equal and opposite as expected.}.
It can be seen that the output is held in a steady state according
to the delta inputs as expected.

% {{{ fig:dc2_plot
\begin{figure}[htbp!]
\includegraphics[scale=0.7]{../octave/dc2_plot}
\caption{Response of delta input to steady state output system
when given an arbitrary input signal.
The upper plot is the input signal ($u$) and the lower plot is
the output response ($y$).
The Matlab source code is given in Listing \ref{lst:dc2_init}
and \ref{lst:dc2_plot}.
}
\label{fig:dc2_plot}
\end{figure}
% }}}

\clearpage
\subsection{Matlab Source}
\label{app:dcsrc}

%\lstinputlisting[caption={Matlab code to initialize the full
%steady state to delta system.},
%	label=lst:dc1_init
%	]{../octave/dc1_init.m}
%
%\lstinputlisting[caption={Matlab code to plot the full steady state
%to delta system.},
%	label=lst:dc1_plot
%	]{../octave/dc1_plot.m}

\lstinputlisting[caption={Matlab code to initialize the simplified
steady state to delta system.},
	label=lst:dc2_init
	]{../octave/dc2_init.m}

\lstinputlisting[caption={Matlab code to plot the simplified steady state
to delta system.},
	label=lst:dc2_plot
	]{../octave/dc2_plot.m}
